Comparison of Terrain Following and Cut Cell Grids with a Non-Hydrostatic Model

### 2015-02-04 msc-publication.tex
reorder intro paras, say what this study is trying to do

It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998} \TODO{or steppeler2002}.  Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.

@@

\TODO{that's the high-level intro, now onto BTF/SLEVE grids...}  A variety of vertical decay functions have been formulated that make TF layers flatter with height.  \citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as \(z = \left( H - h \right) \left( z^\star / H \right) + h\) where, in two dimensions, \(z(x, z^\star)\) is the height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.

The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain.  The terrain height is split into large-scale and small-scale components, \(h_1\) and \(h_2\), such that \(h = h_1 + h_2\), with each component having a different exponential decay. The transformation is defined as \(z = z^\star + h_1 b_1 + h_2 b_2\), where the vertical decay functions are given by
\begin{displaymath}
	b_i = \frac{\sinh \left( \left( H - z^\star \right) / s_i \right)}{\sinh \left( H / s_i \right)}
\end{displaymath}
with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.

\TODO{say what we're going to do in this paper}

\TODO{final para of intro: in section 2, section 3 etc etc}

\begin{itemize}
	\item Comparing results BTF, SLEVE and cut cell style grid for three standard two-dimensional test cases
	\item TF grids often shown to have lesser accuracy than cut cells in some test cases, often significantly so.  Our results demonstrate that accuracy need not be much reduced and can even be better than cut cells. \TODO{this is too confrontational!}
\end{itemize}

\section{Results}
\begin{itemize}
	\item \TODO{How much do I need to say about the model?  -- Give a summary: what equation set? finite volume with curl-free pressure grad; lorenz grid; advection scheme}
	\item \TODO{Explain how SnapCol grid is constructed?}
\end{itemize}

@@

\begin{itemize}
	\item Motivation: challenge accuracy of H operator (\TODO{if we mention H op here we'll have to explain it beforehand})
	\item Specify domain, thermodynamics, discretisation in time and space
\end{itemize}

@@

Analysis:
\begin{itemize}
	\item $w$ contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
	\item \TODO{don't know if divergence is relevant -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
	\item $\theta$ anomalies similar on all grids EXCEPT...
	\item ... on SnapCol grid in lee of mountain near the ground. (Figure~\ref{fig:gw-theta})
	\item \TODO{worth mentioning implications of this?  1. reduced stability, although not enough to create vertical motion in this instance 2. model has no viscosity so thermal mixing should not occur in theory}
	\item Lorenz computation mode has been excited because Exner sample line at $x = \SI{50}{\kilo\meter}$ is in hydrostatic balance (Figure~\ref{fig:gw-exner-theta})
	\item \TODO{We can speculate on what excites the computational mode, but perhaps better not to?}
	\item Little/no evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid
\end{itemize}

>>>>

\citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as 
\begin{equation}
	z = \left( H - h \right) \left( z^\star / H \right) + h
\end{equation}
where, in two dimensions, \(z(x, z^\star)\) is the height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.  Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.

It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998} \TODO{or steppeler2002}.

Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.

The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain than the BTF coordinate.  The terrain height is split into large-scale and small-scale components, \(h_1\) and \(h_2\), such that \(h = h_1 + h_2\), with each component having a different exponential decay. The transformation is defined as 
\begin{align}
	z &= z^\star + h_1 b_1 + h_2 b_2
\intertext{where the vertical decay functions are given by}
	b_i &= \frac{\sinh \left( \left( H - z^\star \right) / s_i \right)}{\sinh \left( H / s_i \right)}
\end{align}
with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.

@@

This study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between BTF, SLEVE and cut cell grids for idealised, two-dimensional test cases from the literature.  \TODO{final bit of intro: in section 2, section 3 etc etc}

\begin{itemize}
	\item TF grids often shown to have lesser accuracy than cut cells in some test cases, often significantly so.  Our results demonstrate that accuracy need not be much reduced and can even be better than cut cells. \TODO{this is too confrontational!}
\end{itemize}

\section{Results}
\begin{itemize}
	\item \TODO{Give a summary of model: fully-compressible, finite volume with curl-free pressure grad; lorenz grid; advection scheme}
	\item \TODO{Explain how SnapCol grid is constructed?}
\end{itemize}

@@

\begin{itemize}
	\item Motivation: challenge accuracy of H operator (\TODO{if we mention H op here we'll have to explain it beforehand}), curl-free pressure gradient formulation
	\item Cite weller-shahrokhi2014 for domain, thermodynamics, discretisation in time and space
\end{itemize}

@@

Analysis:
\begin{itemize}
	\item $w$ contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
	\item \TODO{divergence might help explain computational mode -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
	\item $\theta$ anomalies similar on all grids EXCEPT...
	\item ... on SnapCol grid in lee of mountain near the ground. (Figure~\ref{fig:gw-theta})
	\item \TODO{worth mentioning implications of this?  1. reduced stability, although not enough to create vertical motion in this instance 2. would disrupt clouds 2. model has no viscosity so thermal mixing should not occur in theory}
	\item Lorenz computation mode has been excited because Exner sample line at $x = \SI{50}{\kilo\meter}$ is in hydrostatic balance (Figure~\ref{fig:gw-exner-theta})
	\item \TODO{We can speculate on what excites the computational mode, but perhaps better not to?}
	\item No evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid
\end{itemize}
